Research Notes

### 2019-02-01 writeups/notes.tex
check in some temp paper work

If we approximate $F(x, z, t) \approx F_0\delta\p*{z - z_0} e^{i(k_xx - \omega t)}$
for the time being, then clearly the only term that can be equally singular is
the $\ptd{}{z}$ term, leaving us with
\begin{equation}
    -\omega^2 \ptd{u_{1z}}{z} - \frac{Fgk_x^2}{\rho_0} = 0.
\end{equation}
Since $u_{1z}$ both above and below $z_0$ satisfy the homogeneous PDE, which has
plane wave solutions with an $e^{z/2H}$ envelope where $k_zH \gg 1$, we can
integrate about a small domain around $z_0$ and obtain
\begin{equation}
    \at{\pd{u_{1z}}{z}}_{z > z_0} - \at{\pd{u_{1z}}{z}}_{z < z_0}
        = -F_0\frac{gk_x^2}{\rho_0\omega^2}e^{i(k_xx - \omega t)}.
\end{equation}
It should then be apparent that the $u_{1z}$ solution is
\begin{equation}
    u_{1z} = \frac{F_0gk_x^2}{\rho_0(z = z_0)\omega^2}\frac{1}{2ik_z} \times
    \begin{cases}
        e^{(z - z_0)/2H}e^{i(k_xx + k_z(z - z_0) - \omega t)} & z > z_0\\
        e^{(z - z_0)/2H}e^{i(k_xx - k_z(z - z_0) - \omega t)} & z < z_0\\
    \end{cases}.
\end{equation}

@@

>>>>

If we approximate $F(x, z, t) \approx F_0\delta\p*{z - z_0} e^{i(k_xx - \omega t)}$
for the time being, then clearly the only term that can be equally singular is
the $\ptd{}{z}$ term, leaving us with
\begin{equation}
    -\omega^2 \ptd{u_{1z}}{z} - \frac{Fgk_x^2}{\rho_0} = 0.
\end{equation}
Since $u_{1z}$ both above and below $z_0$ satisfy the homogeneous PDE, which has
plane wave solutions with an $e^{z/2H}$ envelope where $k_zH \gg 1$, we can
integrate about a small domain around $z_0$ and obtain
\begin{equation}
    \at{\pd{u_{1z}}{z}}_{z > z_0} - \at{\pd{u_{1z}}{z}}_{z < z_0}
        = -F_0\frac{gk_x^2}{\rho_0\omega^2}e^{i(k_xx - \omega t)}.
\end{equation}
It should then be apparent that the $u_{1z}$ solution is
\begin{equation}
    u_{1z} = \frac{F_0gk_x^2}{\rho_0(z = z_0)\omega^2}\frac{1}{2ik_z} \times
    \begin{cases}
        e^{(z - z_0)/2H}e^{i(k_xx + k_z(z - z_0) - \omega t)
            + \frac{1}{2k_zH}} & z > z_0\\
        e^{(z - z_0)/2H}e^{i(k_xx - k_z(z - z_0) - \omega t)
            + \frac{1}{2k_zH}} & z < z_0\\
    \end{cases}.
\end{equation}
The extra phase shift comes from $\pd{}{z} \approx ik_z + \frac{1}{2H} \approx
k_z\p*{e^{-\frac{1}{2k_zH}}}$.

@@

\chapter{Tidal Synchronization}

Chapter containing notes slightly more pertaining to the paper. I've kinda not
been updating my notes here for the last few months, so I'll just add what I
can.

\section{Momentum-Conserving Equations}

First, we want to simplate the Euler equations, but $\frac{\vec{\nabla}P}{\rho}$
is difficult to deal with. Instead, define $\varpi = P / \rho$, then we can
divide $\rho$ out of the momentum equation and get no quotients of dynamical
variables. It is then easier to use $\Upsilon: \rho = \rho e^\Upsilon$.
Proceeding:

We start from the flux-conservative Euler equations and manually insert
artificial viscosity terms in a momentum-conserving way:
\begin{subequations}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} &= 0,\\
        \partial_t \rho + \vec{\nabla} \cdot (\rho \vec{u} - \nu
            \vec{\nabla}(\rho - \rho_0)) &= 0,\label{eq:visc_cons_mom}\\
        \partial_t (\rho \vec{u}) + \vec{\nabla} \cdot (\rho \vec{u} \vec{u} +
            \mathrm{diag}(\rho \varpi) - \nu \rho \vec{\nabla}\vec{u}) + \rho g
            \hat{z} &= 0.
    \end{align}
\end{subequations}
We can do some algebra and obtain
\begin{subequations}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} &= 0,\\
        \partial_t \Upsilon + \p*{\vec{u} \cdot \vec{\nabla}} \Upsilon -
            \frac{u_z}{H} - \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla}
            \Upsilon} \cdot \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z
            \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}} &= 0,\\
        \partial_t \vec{u} + \p*{\vec{u} \cdot \vec{\nabla}}\vec{u} +
            \vec{\nabla} \varpi + \varpi \vec{\nabla} \Upsilon - \nu \nabla^2
            \vec{u} + \vec{u} \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla}
            \Upsilon} \cdot \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z
            \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}}&{}\nonumber\\
        - 2\nu \p*{\p*{\p*{\vec{\nabla}\Upsilon} \cdot \vec{\nabla}}\vec{u} -
            \frac{1}{H}\partial_z \vec{u}} - \frac{\varpi_1}{H} &= 0.
    \end{align}
\end{subequations}
Including numerical dissipation, sponge layers and driving gives the full system
of equations
\begin{subequations}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} ={}& 0,\\
        \partial_t \Upsilon - \frac{u_z}{H}
            ={}& \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla}
            \Upsilon} \cdot \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z
            \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}},\nonumber\\
            & - \p*{\vec{u} \cdot \vec{\nabla}}\Upsilon
                -\Gamma(z) \Upsilon
                + \frac{F}{\rho_0(z)}e^{-\frac{(z - z_0)^2}{2\sigma^2}}
                    \cos \p*{k_xx - \omega t},\\
        \pd{u_x}{t} + \pd{T}{x} + gH\pd{\Upsilon}{x} ={}&
            \nu \nabla^2 u_x
            - u_x \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla} \Upsilon} \cdot
                \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z \Upsilon
                + \frac{1 - e^{-\Upsilon}}{H^2}}\nonumber\\
            &+ 2\nu \p*{\p*{\p*{\vec{\nabla}\Upsilon} \cdot \vec{\nabla}}u_x
                - \frac{1}{H}\partial_z u_x}
            -\Gamma(z) u_x
                - \p*{\vec{u} \cdot \vec{\nabla}}u_x
                - T_1 \pd{\Upsilon}{x},\\
        \pd{u_z}{t} + \pd{T}{z} + gH\pd{\Upsilon}{z} - \frac{T_1}{H} ={}&
            \nu \nabla^2 u_z
            - u_z \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla} \Upsilon} \cdot
                \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z \Upsilon
                + \frac{1 - e^{-\Upsilon}}{H^2}}\nonumber\\
            &+ 2\nu \p*{\p*{\p*{\vec{\nabla}\Upsilon} \cdot \vec{\nabla}}u_z -
                \frac{1}{H}\partial_z u_{z}}
            -\Gamma(z) u_z - \p*{\vec{u} \cdot \vec{\nabla}}u_z
            - T_1 \pd{\Upsilon}{z},\\
        \Gamma(z) &= 7.5\s*{2 + \tanh \frac{z - z_T}{(L_z - z_T) / 2}
            + \tanh \frac{z_B - z}{z_B / 2}},\label{eq:Gamma}
    \end{align}
\end{subequations}

\section{Mean Flow Absorption}

The model we're examining now is if part of the flow has synchronized to the
horizontal phase velocity of the wave. Then as the wavetrain is incident on the
wave with horizontal momentum flux $S_{px}$, it completely deposits its
horizonttal momentum. Thus, $S_{px}$ goes from the wave's, which is well known,
down to zero, being deposited into the mean flow, which forces
\begin{equation}
    \rho(z) c_{ph,x} \pd{z_c}{t} = -S_{px}.
\end{equation}
where $c_{ph, x}$ is the horizontal phase velocity/critical velocity, $z_c$ is
the location of the dissipation front, and $S_{px}$ constant is the incident
momentum flux. This can be analytically solved if we take $\rho(z) \approx
\rho_0 \propto e^{-z/H}$. Try ansatz $z(t) = -H\p*{\ln t + \ln \tau}$ and we
find
\begin{equation}
    z(t) = -H\ln t - H\ln \frac{H\rho_0(z = 0)c_{ph, x}}{S_{px}}.
\end{equation}
